Globally-centered autocovariances in MCMC

### 2021-07-21 Paper/Submissions/JCGS/First/arxiv/main.tex

\onehalfspacing
\begin{abstract}
Autocovariances are a fundamental quantity of interest in Markov chain Monte Carlo (MCMC) simulations with autocorrelation function (ACF) plots being an integral visualization tool for performance assessment. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple-chain MCMC sampling,  we propose a globally-centered estimator of the autocovariance function (G-ACvF) that exhibits significant theoretical and empirical improvements. We show that the bias of the G-ACvF estimator is smaller than the bias of the current state-of-the-art.  The impact of this improved estimator is evident in three critical output analysis applications: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. Under weak conditions, we establish strong consistency of an improved asymptotic covariance estimator, and obtain its large-sample bias and variance.  The performance of the new estimators is demonstrated through various examples.
\end{abstract}

@@

We demonstrate the striking difference in estimation of $\Gamma(k)$ via ACF plots. Consider a random-walk Metropolis-Hastings sampler for a univariate mixture of Gaussians target density. Let
\[
f(x) = 0.7\,f(x; -5, 1) + 0.3\,f(x; 5, 0.5)\,,
\]
be the target density where $f(x; a,b)$ is the density of a normal distribution with mean $a$ and variance $b$. 
We set $m = 2$ with starting values distributed to the two modes. The trace plots in Figure~\ref{fig:gaussian-trace} indicate that in the first $10^4$ steps, the chains do not jump modes so that both Markov chains yield significantly different estimates of the population mean, $\mu$. At $10^5$ sample size, both Markov chains have traversed the state space reasonably and yield similar estimates of $\mu$.

@@

\begin{figure}[htbp]
\centering
   \includegraphics[width=.62\linewidth]{plots/gaussian-Targettrace.pdf}\\
\vspace{-.5cm}
    \includegraphics[width=.62\textwidth]{plots/gaussian-acf_hist.pdf} 
    \caption{Top: Target density and trace plots for two chains at $n = 10^4$ (left) and $n = 10^5$ (right). Bottom: ACF plots for the first chain using local-centering (left) and global-centering (right). Histogram estimates at $n = 10^4$ with the blue curve being estimates from $n = 10^5$.}
    \label{fig:gaussian-trace}
\end{figure}

% \begin{figure}[htbp]
% \centering
% \includegraphics[width=.6\textwidth]{plots/gaussian-acf_hist.pdf} \\
% % \includegraphics[width=.6\textwidth]{plots/gaussian-acf_1e4.pdf}

@@

We end this section by noting that we can obtain an average G-ACvF and G-ACF over all $m$ chains
\[
\hat{\Gamma}_G(k) = \dfrac{1}{m}\sum_{s=1}^m \hat{\Gamma}_{G,s}(k) \quad \text{ and } \quad \hat{\rho}^{(i)}_G(k) = \dfrac{1}{m}\sum_{s=1}^m \hat{\rho}^{(i)}_{G,s}(k)\,.
\]
The averaged estimators present a measure of the overall correlation structure induced by the Markov transition $P$.

@@

% \end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$ \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. There is a rich literature on estimating $\Sigma$ for single-chain MCMC implementations. The most common are SV estimators \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from parallel chains. Batch means estimators are computationally more efficient than SV estimators, whereas SV estimators are more reliable \citep{damerdji:1995,fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV (G-SV) estimator of $\Sigma$. Using the method of \cite{heberle2017fast}, we also provide a computationally efficient implementation of the G-SV estimator.

@@

\begin{theorem}[\cite{kuel:1976,vats:fleg:jon:2018}]
  \label{thm:kuelbs}
Let $\E_F\|Y_{11}\|^{2+ \delta} < \infty$ for $\delta > 0$ and let $P$ be polynomially ergodic of order $\xi > (q + 1 + \epsilon)/(1 + 2/\delta)$ for $q \geq 1$. Then there exists a $p \times p$ lower triangular matrix $L$ with $LL^T = \Sigma$, a non-negative function $\psi(n) = n^{1/2 - \lambda}$ for some $\lambda > 0$, a finite random variable $D$, and a sufficiently rich probability space $\Omega$ such that for all $n > n_0$,
\[
\left\|\sum_{t=1}^{n}Y_t - n\mu_g - LB(n)\right\| < D\psi(n) \qquad \text{  with probability 1}\,.
\]
\end{theorem}

@@

\begin{algorithm}[htbp] 
\KwIn{$C(w^*)$ and $A_s$}
\DontPrintSemicolon
\SetAlgoLined
% Construct $C(w^*) \textrm{ and } B_s^*$ from $\{Y_{st};\, t = 1, ...,\, n\}$.\\ 
Compute eigenvalues $\lambda_i$ of  $C(w^*)_{(1)}$ using a discrete Fourier transform, $i = 1, \dots, 2n$\;
Construct $2n \times p$ matrix $A^*_s = (A_s^T  \quad 0_{n \times q})^T$\;

@@

For three different target distributions we sample $m$ parallel Markov chains to assess the performance of our proposed estimators. We make the following three comparisons - (1) locally-centered ACF vs G-ACF estimator, (2) A-SV vs G-SV estimator, and (3) $\widehat{\textrm{ESS}}_A$ vs $\widehat{\textrm{ESS}}_G$. The quality of estimation of $\Sigma$ is studied by coverage probabilities of a $95 \%$ Wald  confidence region when the true mean $\mu_g$ is known. The convergence of local and global estimators of $\Sigma$ and $\widehat{\textrm{ESS}}$ as $n$ increases is studied through two types of running plots (1) logarithm of Frobenius norm of estimated $\Sigma$, and (2) logarithm of estimated ESS/$mn$. In all three examples, we estimate the mean of the stationary distribution, so $g$ is the identity function. The truncation point $b_n$ are the defaults available in the \texttt{R} package \texttt{mcmcse} \citep{mcmcse}.

@@

\begin{figure}[htbp]
\centering
   \includegraphics[width=.5\linewidth]{plots/var-acf_n1000.pdf}
     \includegraphics[width=.5\linewidth]{plots/var-acf_n10000.pdf} 
\caption{VAR. ACF (left) and G-ACF (right) for the second chain for $m = 5$. (Top) $n = 10^3$  and (bottom) $n =  10^4$. The red line is the true ACF.}
\label{fig:var-acf}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = 2.0in]{plots/var-frob.pdf} 
      \includegraphics[width = 2.0in]{plots/var-ess.pdf}         
    \caption{VAR. (Left) Running plot for logarithm of Frobenius norm of A-SV and G-SV estimator. (Right) Running plot for logarithm of $\widehat{\textrm{ESS}}/mn$ using A-SV and G-SV estimator.}
    \label{fig:var-frob_n_ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .70\textwidth]{plots/boom-all_1_3_8.pdf}\\
    \includegraphics[width = .70\textwidth]{plots/boom-all_1_10_7.pdf}    
    \caption{Boomerang. Top row is Setting 1 and bottom row is Setting 2. Contour plots of the target distributions  overlaid with scatter plots for two chains.  Locally-centered ACF (middle) and G-ACF (right) plots for one chain for $n = 10^3$ with the blue line being the ACF at $n = 10^4$.  }
   \label{fig:boom-2D}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .30\textwidth]{plots/boom-ess_1_3_8_m5.pdf}
      \includegraphics[width = .30\textwidth]{plots/boom-ess_1_10_7_m5.pdf}

    \caption{Boom: Running plot of $\log (\widehat{\text{ESS}}_A )/mn$ and $\log (\widehat{\text{ESS}}_G )/mn$ with $m = 5$ for setting 1 (left) and setting 2 (right).}
    \label{fig:boom-ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .32\textwidth]{plots/sensor-trace_loc1.pdf}
      \includegraphics[width = .32\textwidth]{plots/sensor-acf_n5e3.pdf}
      \includegraphics[width = .32\textwidth]{plots/sensor-acf_n5e4.pdf}    
    \caption{Sensor: Trace plot of $x_{11}$ (left) for two parallel chains. Average locally-centered ACF (solid orange) and G-ACF (solid blue) at $n = 5000$ (middle) and $n = 50000$ (right). Dashed lines are individual  chain estimates.}
    \label{fig:sensor-trace}
\end{figure}

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .30\textwidth]{plots/sensor-frob.pdf} 
      \includegraphics[width = .30\textwidth]{plots/sensor-ess.pdf}

@@

Another class of estimators for $\Sigma$ for reversible Markov chains are the multivariate initial sequence (mIS) estimators \citep{dai:jon:2017}. Similar to SV estimators, ACvFs are a critical part of mIS estimators and underestimation in ACvFs yields underestimates of $\Sigma$. It is easy to show that $\hat{\Upsilon}_{G,s}(k)$ is a strongly consistent estimators for $\Upsilon(k)$. Replacing $\hat{\Upsilon}_s(k)$ for $\hat{\Upsilon}_{G,s}(k)$ and following \cite[Theorem~2]{dai:jon:2017} will yield consistent overestimation of the generalized variance $\abs{\Sigma}$ estimated using a globally-centered version of the mIS estimator.  A valuable line of research would be to study the theoretical and experimental analysis of mIS estimators using G-ACvF.

@@

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Trace plot for second (left) and third (right) component of two chains.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF plots using local (left) and global (right) centering for $m=2$ parallel Markov chains. The individual chains are shown through dashed lines and average over $m$ chains is the solid line. The chain lengths are $n = 1000$ (top row) and $n = 10000$ (bottom row).}
    \label{fig:poisson-acf}
\end{figure}

@@

The goal is to draw nodes uniformly from the network by using a network crawler. \cite{nilakanta2019ensuring} modified the data by removing 1,022 out of 1,461 nodes to obtain a well-connected graph. This resulting social network has 439 nodes and 573 edges. We use a Metropolis-Hastings algorithm with a simple random-walk  proposal suggested by \cite{gjoka2011practical}. Each node is associated with five features namely - degree of connection, cluster coefficient, grade, binary sex indicator (1 for female, 0 for male), and binary race indicator (1 for white, 0 for others)

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/magnolia-acf_n100.pdf}\\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/magnolia-acf_n1000.pdf}

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network. 
We sample two parallel Markov chains starting from two students belonging to different races.
 % and study its impacts on the average features of their immediate social group, especially their grades. Due to cluster formation in the network, a Markov chain will land on students within a cluster more often than on students outside it. As a consequence, the sample averages for the two chains will differ for a relatively shorter chain length. 
% 
Figure~\ref{fig:magnolia-acf} shows the ACF plots for the \textit{grade} feature at two different simulation sizes for $m=2$ parallel Markov chains. For a shorter chain length, both the chains have explored different clusters and as a consequence, the local and global mean do not agree. Regardless, G-ACF displays a clear advantage over locally-centered ACF for a chain length of $n=100$ samples whereas the latter takes $n=1000$ samples to reach the truth.

% It is hypothesized that students from certain communities tend to engage more with students of same or other marginally represented communities. We believe that this "selective networking" might cause formation of clusters in the network wherein students within a cluster engage more with each other than with the students outside it. Given that the graph is well-connected, the Markov chains will eventually explore all the nodes.

\singlespacing
\bibliographystyle{apalike}
\bibliography{sample}
\end{document}

>>>>

\doublespacing
\begin{abstract}
Autocovariances are a fundamental quantity of interest in Markov chain Monte Carlo (MCMC) simulations with autocorrelation function (ACF) plots being an integral visualization tool for performance assessment. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple-chain MCMC sampling,  we propose a globally-centered estimator of the autocovariance function (G-ACvF) that exhibits significant theoretical and empirical improvements. We show that the bias of the G-ACvF estimator is smaller than the bias of the current state-of-the-art.  The impact of this improved estimator is evident in three critical output analysis applications: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. Under weak conditions, we establish strong consistency of our improved asymptotic covariance estimator, and obtain its large-sample bias and variance.  The performance of the new estimators is demonstrated through various examples.
\end{abstract}

@@

We use ACF plots to demonstrate the striking difference in the estimation of $\Gamma(k)$. Consider a random-walk Metropolis-Hastings sampler for a univariate mixture of Gaussians target density. Let
\[
f(x) = 0.7\,f(x; -5, 1) + 0.3\,f(x; 5, 0.5)\,,
\]
be the target density where $f(x; a,b)$ is the density of a normal distribution with mean $a$ and variance $b$. 
We set $m = 2$ with starting values distributed to the two modes. The trace plots in Figure~\ref{fig:gaussian-trace} indicate that in the first $10^4$ steps, the chains do not jump modes so that both Markov chains yield significantly different estimates of the population mean, $\mu$. At $10^5$ sample size, both Markov chains have traversed the state space reasonably and yield similar estimates of $\mu$.

@@

\begin{figure}[htbp]
\centering
   \includegraphics[width=.62\linewidth]{gaussian-TargetTrace.pdf}\\
\vspace{-.5cm}
    \includegraphics[width=.62\textwidth]{gaussian-acf_hist.pdf} 
    \caption{Top: Target density and trace plots for two chains at $n = 10^4$ (left) and $n = 10^5$ (right). Bottom: ACF plots for the first chain using local-centering (left) and global-centering (right). Histogram estimates at $n = 10^4$ with the blue curve being estimates from $n = 10^5$.}
    \label{fig:gaussian-trace}
\end{figure}

% \begin{figure}[htbp]
% \centering
% \includegraphics[width=.6\textwidth]{gaussian-acf_hist.pdf} \\
% % \includegraphics[width=.6\textwidth]{gaussian-acf_1e4.pdf}

@@

We end this section by noting that we can obtain an average G-ACvF and G-ACF over all $m$ chains as,
\[
\hat{\Gamma}_G(k) = \dfrac{1}{m}\sum_{s=1}^m \hat{\Gamma}_{G,s}(k) \quad \text{ and } \quad \hat{\rho}^{(i)}_G(k) = \dfrac{1}{m}\sum_{s=1}^m \hat{\rho}^{(i)}_{G,s}(k)\,.
\]
The averaged estimators present a measure of the overall correlation structure induced by the Markov transition $P$.

@@

% \end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$ \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. There is a rich literature on estimating $\Sigma$ for single-chain MCMC implementations. The most common are SV estimators \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from parallel Markov chains. Batch means estimators are computationally more efficient than SV estimators, whereas SV estimators are more reliable \citep{damerdji:1995,fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV (G-SV) estimator of $\Sigma$. Using the method of \cite{heberle2017fast}, we also provide a computationally efficient implementation of the G-SV estimator.

@@

\begin{theorem}[\cite{kuel:phil:1980,vats:fleg:jon:2018}]
  \label{thm:kuelbs}
Let $\E_F\|Y_{11}\|^{2+ \delta} < \infty$ for $\delta > 0$ and let $P$ be polynomially ergodic of order $\xi > (q + 1 + \epsilon)/(1 + 2/\delta)$ for $q \geq 1$. Then there exists a $p \times p$ lower triangular matrix $L$ with $LL^T = \Sigma$, a non-negative function $\psi(n) = n^{1/2 - \lambda}$ for some $\lambda > 0$, a finite random variable $D$, and a sufficiently rich probability space $\Omega$ such that for all $n > n_0$,
\[
\left\|\sum_{t=1}^{n}Y_t - n\mu_g - LB(n)\right\| < D\psi(n) \qquad \text{  with probability 1}\,.
\]
\end{theorem}

@@

\begin{algorithm}[htbp] 
\KwIn{$C(w^*)$ and $A_s$}
\DontPrintSemicolon
\SetAlgoLined
% Construct $C(w^*) \textrm{ and } B_s^*$ from $\{Y_{st};\, t = 1, ...,\, n\}$.\\ 
Compute eigenvalues $\lambda_i$ of  $C(w^*)_{(1)}$ using a discrete Fourier transform, $i = 1, \dots, 2n$\;
Construct $2n \times p$ matrix $A^*_s = (A_s^T  \quad 0_{n \times p})^T$\;

@@

For three different target distributions we sample $m$ parallel Markov chains to assess the performance of our proposed estimators. We make the following three comparisons - (1) locally-centered ACF vs G-ACF estimator, (2) A-SV vs G-SV estimator, and (3) $\widehat{\textrm{ESS}}_A$ vs $\widehat{\textrm{ESS}}_G$. The quality of estimation of $\Sigma$ is studied by coverage probabilities of a $95 \%$ Wald  confidence region when the true mean $\mu_g$ is known. The convergence of local and global estimators of $\Sigma$ and $\textrm{ESS}$ as $n$ increases is studied through two types of running plots (1) logarithm of Frobenius norm of estimated $\Sigma$, and (2) logarithm of estimated ESS/$mn$. In all three examples, we estimate the mean of the stationary distribution, so $g$ is the identity function. The truncation point $b_n$ are the defaults available in the \texttt{R} package \texttt{mcmcse} \citep{mcmcse}.

@@

\begin{figure}[htbp]
\centering
   \includegraphics[width=.60\linewidth]{var-acf_n1000.pdf}
     \includegraphics[width=.60\linewidth]{var-acf_n10000.pdf} 
\caption{VAR. ACF (left) and G-ACF (right) for the second chain for $m = 5$. (Top) $n = 10^3$  and (bottom) $n =  10^4$. The red line is the true ACF.}
\label{fig:var-acf}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = 2.3in]{var-frob.pdf} 
      \includegraphics[width = 2.3in]{var-ess.pdf}         
    \caption{VAR. (Left) Running plot for logarithm of Frobenius norm of A-SV and G-SV estimator. (Right) Running plot for logarithm of $\widehat{\textrm{ESS}}/mn$ using A-SV and G-SV estimator.}
    \label{fig:var-frob_n_ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .70\textwidth]{boom-all_1_3_8.pdf}\\
    \includegraphics[width = .70\textwidth]{boom-all_1_10_7.pdf}    
    \caption{Boomerang. Top row is Setting 1 and bottom row is Setting 2. Contour plots of the target distributions  overlaid with scatter plots for two chains.  Locally-centered ACF (middle) and G-ACF (right) plots for one chain for $n = 10^3$ with the blue line being the ACF at $n = 10^4$.  }
   \label{fig:boom-2D}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .30\textwidth]{boom-ess_1_3_8_m5.pdf}
      \includegraphics[width = .30\textwidth]{boom-ess_1_10_7_m5.pdf}

    \caption{Boomerang. Running plot of $\log (\widehat{\text{ESS}}_A )/mn$ and $\log (\widehat{\text{ESS}}_G )/mn$ with $m = 5$ for setting 1 (left) and setting 2 (right).}
    \label{fig:boom-ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .32\textwidth]{sensor-trace_loc1.pdf}
      \includegraphics[width = .32\textwidth]{sensor-acf_n5e3.pdf}
      \includegraphics[width = .32\textwidth]{sensor-acf_n5e4.pdf}    
    \caption{Sensor: Trace plot of $x_{11}$ (left) for two parallel chains. Average locally-centered ACF (solid orange) and G-ACF (solid blue) at $n = 5000$ (middle) and $n = 50000$ (right). Dashed lines are individual  chain estimates.}
    \label{fig:sensor-trace}
\end{figure}

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .30\textwidth]{sensor-frob.pdf} 
      \includegraphics[width = .30\textwidth]{sensor-ess.pdf}

@@

Another class of estimators for $\Sigma$ for reversible Markov chains are the multivariate initial sequence (mIS) estimators \citep{dai:jon:2017}. Similar to SV estimators, ACvFs are a critical part of mIS estimators and underestimation in ACvFs yields underestimates of $\Sigma$. It is easy to show that $\hat{\Upsilon}_{G,s}(k)$ is a strongly consistent estimators for $\Upsilon(k)$. Replacing $\hat{\Upsilon}_s(k)$ for $\hat{\Upsilon}_{G,s}(k)$ and following \cite[Theorem~2]{dai:jon:2017} will yield consistent overestimation of the generalized variance $\abs{\Sigma}$ resulting in a globally-centered version of the mIS estimator.  A valuable line of research would be to study the theoretical and empirical properties of the globally-centered mIS estimators.

@@

% \bigskip
% \begin{center}
% {\large\bf SUPPLEMENTARY MATERIAL}
% \end{center}

% \begin{description}

% \item[R Code:] Contains R code that reproduces the simulations, figured, and tables (.zip file).

% \end{description}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{poisson-trace_n50000.pdf}
    \caption{Change point: Trace plots for second (left) and third (right) component.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{poisson-acf_n10000.pdf}      
    \caption{Change point: ACF plots using local (left) and global (right) centering for $m=2$ parallel Markov chains. The individual chains are shown through dashed lines and average over $m$ chains is the solid line. The chain lengths are $n = 1000$ (top row) and $n = 10000$ (bottom row).}
    \label{fig:poisson-acf}
\end{figure}

@@

The goal is to draw nodes uniformly from the network by using a network crawler. \cite{nilakanta2019ensuring} modified the data by removing 1,022 out of 1,461 nodes to obtain a well-connected graph. This resulting social network has 439 nodes and 573 edges. We use a Metropolis-Hastings algorithm with a simple random-walk  proposal suggested by \cite{gjoka2011practical}. Each node is associated with five features namely - degree of connection, cluster coefficient, grade, binary sex indicator (1 for female, 0 for male), and binary race indicator (1 for white, 0 for others). We sample two parallel Markov chains starting from two students belonging to different races.

\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{magnolia-acf_n100.pdf}\\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{magnolia-acf_n1000.pdf}

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network.

% 
Figure~\ref{fig:magnolia-acf} shows the ACF plots for the \textit{grade} feature at two different simulation sizes for $m=2$ parallel Markov chains. For a shorter chain length, both the chains have explored different clusters and as a consequence, the local and global mean do not agree. Regardless, G-ACF displays a clear advantage over locally-centered ACF for a chain length of $n=100$ samples whereas the latter takes $n=1000$ samples to reach the truth.

\bibliographystyle{apalike}
\bibliography{sample}
\end{document}

### 2021-07-21 Paper/Submissions/JCGS/First/main.tex

\bigskip
\begin{abstract}
Autocovariances are a fundamental quantity of interest in Markov chain Monte Carlo (MCMC) simulations with autocorrelation function (ACF) plots being an integral visualization tool for performance assessment. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple-chain MCMC sampling,  we propose a globally-centered estimator of the autocovariance function (G-ACvF) that exhibits significant theoretical and empirical improvements. We show that the bias of the G-ACvF estimator is smaller than the bias of the current state-of-the-art.  The impact of this improved estimator is evident in three critical output analysis applications: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. Under weak conditions, we establish strong consistency of an improved asymptotic covariance estimator, and obtain its large-sample bias and variance.  The performance of the new estimators is demonstrated through various examples.
\end{abstract}

@@

    \caption{Boom: Running plot of $\log (\widehat{\text{ESS}}_A )/mn$ and $\log (\widehat{\text{ESS}}_G )/mn$ with $m = 5$ for setting 1 (left) and setting 2 (right).}
    \label{fig:boom-ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Trace plot for second (left) and third (right) component of two chains.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF plots using local (left) and global (right) centering for $m=2$ parallel Markov chains. The individual chains are shown through dashed lines and average over $m$ chains is the solid line. The chain lengths are $n = 1000$ (top row) and $n = 10000$ (bottom row).}
    \label{fig:poisson-acf}
\end{figure}

@@

The goal is to draw nodes uniformly from the network by using a network crawler. \cite{nilakanta2019ensuring} modified the data by removing 1,022 out of 1,461 nodes to obtain a well-connected graph. This resulting social network has 439 nodes and 573 edges. We use a Metropolis-Hastings algorithm with a simple random-walk  proposal suggested by \cite{gjoka2011practical}. Each node is associated with five features namely - degree of connection, cluster coefficient, grade, binary sex indicator (1 for female, 0 for male), and binary race indicator (1 for white, 0 for others)

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network. 
We sample two parallel Markov chains starting from two students belonging to different races.

>>>>

\bigskip
\begin{abstract}
Autocovariances are a fundamental quantity of interest in Markov chain Monte Carlo (MCMC) simulations with autocorrelation function (ACF) plots being an integral visualization tool for performance assessment. Unfortunately, for slow mixing Markov chains, the empirical autocovariance can highly underestimate the truth. For multiple-chain MCMC sampling,  we propose a globally-centered estimator of the autocovariance function (G-ACvF) that exhibits significant theoretical and empirical improvements. We show that the bias of the G-ACvF estimator is smaller than the bias of the current state-of-the-art.  The impact of this improved estimator is evident in three critical output analysis applications: (1) ACF plots, (2) estimates of the Monte Carlo asymptotic covariance matrix, and (3) estimates of the effective sample size. Under weak conditions, we establish strong consistency of our improved asymptotic covariance estimator, and obtain its large-sample bias and variance.  The performance of the new estimators is demonstrated through various examples.
\end{abstract}

@@

    \caption{Boomerang. Running plot of $\log (\widehat{\text{ESS}}_A )/mn$ and $\log (\widehat{\text{ESS}}_G )/mn$ with $m = 5$ for setting 1 (left) and setting 2 (right).}
    \label{fig:boom-ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = 0.7\textwidth]{plots/poisson-trace_n50000.pdf}
    \caption{Change point: Trace plots for second (left) and third (right) component.}
    \label{fig:poisson-trace}
\end{figure}
\begin{figure}[htbp]
    \centering
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n1000.pdf}  \\ \vspace{-.5cm}
      \includegraphics[width = .6\textwidth]{plots/poisson-acf_n10000.pdf}      
    \caption{Change point: ACF plots using local (left) and global (right) centering for $m=2$ parallel Markov chains. The individual chains are shown through dashed lines and average over $m$ chains is the solid line. The chain lengths are $n = 1000$ (top row) and $n = 10000$ (bottom row).}
    \label{fig:poisson-acf}
\end{figure}

@@

The goal is to draw nodes uniformly from the network by using a network crawler. \cite{nilakanta2019ensuring} modified the data by removing 1,022 out of 1,461 nodes to obtain a well-connected graph. This resulting social network has 439 nodes and 573 edges. We use a Metropolis-Hastings algorithm with a simple random-walk  proposal suggested by \cite{gjoka2011practical}. Each node is associated with five features namely - degree of connection, cluster coefficient, grade, binary sex indicator (1 for female, 0 for male), and binary race indicator (1 for white, 0 for others). We sample two parallel Markov chains starting from two students belonging to different races.

@@

% We believe that the students from different races experience different socio-political conditions which causes them to engage in a "selective networking". This might cause formation of clusters in the social network.
